---
title: "Getting Started with calcbenchR"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with calcbenchR}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
# Only run API-calling chunks if credentials are available
has_creds <- nchar(Sys.getenv("CALCBENCH_USERNAME")) > 0 &&
             nchar(Sys.getenv("CALCBENCH_PASSWORD")) > 0
```

## Introduction

[Calcbench](https://www.calcbench.com) is a financial data service that provides
standardized and as-reported XBRL data from SEC filings. **calcbenchR** is an R
wrapper around the Calcbench REST API, giving you access to:

| Function | Description | Requires Premium |
|---|---|---|
| `get_available_metrics()` | Full metric taxonomy (face, ratio, footnote tables) | No |
| `get_available_companies()` | Company universe with metadata | No |
| `get_filings()` | SEC filing metadata (10-K, 10-Q, 8-K, etc.) | No |
| `get_standardized_data()` | Normalized financial metrics | **Yes** |
| `get_dimensional_data()` | Segment / dimensional data | **Yes** |
| `get_disclosures()` | Footnotes, MD&A, and text block search | No |
| `get_face_statement()` | As-reported financial statements | No |
| `get_business_combinations()` | M&A purchase price allocations | No |
| `get_raw_xbrl()` | Raw as-reported XBRL data points | No |
| `get_non_xbrl_data()` | Non-XBRL numeric data from press releases | No |
| `get_xbrl_tag_values()` | Multi-company XBRL tag / metric values | No |
| `get_13f_data()` | Form 13F institutional holdings data | No |

## Setup

Install from GitHub if you haven't already:

```{r install, eval=FALSE}
# install.packages("devtools")
devtools::install_github("luceydav/calcbenchR")
```

```{r load, eval=has_creds}
library(calcbenchR)
library(data.table)
```

All functions authenticate using environment variables. Set these once per
session (or add them to your `.Renviron` for persistence):

```{r credentials, eval=FALSE}
Sys.setenv(
  CALCBENCH_USERNAME = "your_email@example.com",
  CALCBENCH_PASSWORD = "your_password"
)
```

You can verify they are set:

```{r check-creds, eval=FALSE}
nchar(Sys.getenv("CALCBENCH_USERNAME")) > 0  # should be TRUE
```

> **Tip:** Add `CALCBENCH_USERNAME=...` and `CALCBENCH_PASSWORD=...` to your
> `~/.Renviron` file so credentials are available automatically in every R session.
> Run `usethis::edit_r_environ()` to open the file.

---

## Discover Available Metrics

`get_available_metrics()` returns the full Calcbench metric taxonomy with no
arguments. The result is a named list of three `data.table`s:

| Element     | Contents                                              |
|-------------|-------------------------------------------------------|
| `face`      | Income statement, balance sheet, and cash flow items  |
| `ratio`     | Derived ratios (margins, turns, coverage, etc.)       |
| `footnotes` | Footnote disclosures (lease obligations, segment data, etc.) |

```{r get-metrics, eval=has_creds}
metrics <- get_available_metrics()
names(metrics)
```

Each table has the same structure:

```{r metrics-structure, eval=has_creds}
str(metrics$face[1:3])
```

### Browsing face statement metrics

```{r face-metrics, eval=has_creds}
# First few rows
metrics$face[1:6, .(metric, name, section, period_type)]
```

```{r face-filter, eval=has_creds}
# Find all revenue-related items
metrics$face[grepl("revenue", name, ignore.case = TRUE), .(metric, name, section)]
```

```{r face-sections, eval=has_creds}
# How many metrics per section?
metrics$face[, .N, by = section][order(-N)]
```

### Browsing ratio metrics

```{r ratio-metrics, eval=has_creds}
# Margin-related ratios
metrics$ratio[grepl("margin", name, ignore.case = TRUE), .(metric, name, description)]
```

### Browsing footnote metrics

```{r footnote-metrics, eval=has_creds}
# First few footnote metrics
metrics$footnotes[1:5, .(metric, name, section)]
```

---

## Discover Available Companies

`get_available_companies()` returns a `data.table` of the companies in the
Calcbench universe. Pass an index name to filter by index membership (e.g.
`"SP500"`, `"DJIA"`, `"Russell3000"`), or `NULL` for the full universe.

```{r get-companies, eval=has_creds}
# S&P 500 companies
companies <- get_available_companies(index = "SP500")
nrow(companies)
```

```{r companies-cols, eval=has_creds}
# All available columns
names(companies)
```

```{r companies-head, eval=has_creds}
# Core identifying fields (always populated)
companies[1:5, .(ticker, entity_name, entity_id, CIK)]
```

### Filtering companies

```{r companies-ticker, eval=has_creds}
# Look up a specific ticker
companies[ticker == "MSFT", .(ticker, entity_name, CIK, entity_id)]
```

```{r companies-recent, eval=has_creds}
# Companies with the most recent filings
companies[order(-most_recent_filing_cy, -most_recent_filing_cp),
          .(ticker, entity_name, most_recent_filing_cy, most_recent_filing_cp)][1:10]
```

---

## Get Filings

`get_filings()` retrieves SEC filing metadata for one or more companies. You
can filter by filing type, date range, and whether to include XBRL, non-XBRL,
or press release filings.

```{r get-filings, eval=has_creds}
# Recent filings for Apple (10-Ks and 10-Qs)
aapl_filings <- get_filings(companies = list("AAPL"))
```

```{r filings-preview, eval=has_creds}
# Preview results
head(aapl_filings[, .(ticker, entity_name, filing_type, filing_date, fiscal_year, fiscal_period)])
```

```{r filings-types, eval=has_creds}
# Distribution of filing types
aapl_filings[, .N, by = filing_type][order(-N)]
```

```{r filings-annual, eval=has_creds}
# Annual and quarterly reports only
aapl_filings[filing_type == "annualQuarterlyReport",
             .(ticker, filing_date, fiscal_year, fiscal_period, sec_html_url)][1:5]
```

---

## Get Standardized Financial Data *(Premium)*

> **Note:** `get_standardized_data()` requires a **Calcbench premium subscription**.
> Free and trial accounts will receive a 401 error with a message pointing to
> [calcbench.com/home/plans](https://www.calcbench.com/home/plans).
> The code below is shown with `eval=FALSE`; substitute your own credentials
> and run it interactively if you have a premium account.

`get_standardized_data()` pulls normalized financial metrics across one or more
companies, returning a tidy `data.table` with one row per company/period/metric.

```{r get-standardized, eval=FALSE}
# EBIT for 3M Co, full history
m3_ebit <- get_standardized_data(
  metrics     = list("EBIT"),
  companies   = list("3M Co"),
  all_history = TRUE
)
```

The returned `data.table` looks like this (example output):

```
   calcbench_entity_id ticker fiscal_year fiscal_period metric     value
                 <int> <char>       <int>         <int> <char>     <num>
1:               12345    MMM        2023             0   EBIT  3850000000
2:               12345    MMM        2022             0   EBIT  4100000000
3:               12345    MMM        2021             0   EBIT  3500000000
```

### Pulling multiple metrics and companies

```{r multi-metrics, eval=FALSE}
# Revenue and EBIT for several companies, last 5 years
income_data <- get_standardized_data(
  metrics   = list("Revenue", "EBIT", "NetIncome"),
  companies = list("AAPL", "MSFT", "GOOGL"),
  all_history = FALSE
)
```

### data.table manipulation

Once you have the data, standard `data.table` operations apply:

```{r dt-filter, eval=FALSE}
# Filter to a single metric
revenue <- income_data[metric == "Revenue"]

# Sort by year descending
revenue[order(-fiscal_year, ticker)]

# Wide format: one column per metric
wide <- dcast(income_data, ticker + fiscal_year ~ metric, value.var = "value")
wide[order(ticker, -fiscal_year)]
```

---

## Get Dimensional / Segment Data *(Premium)*

> **Note:** Requires a **Calcbench premium subscription**.

`get_dimensional_data()` retrieves segment-level and other dimensional breakouts
(the data behind [calcbench.com/segments](https://www.calcbench.com/segments)).

```{r get-dimensional, eval=FALSE}
# Operating segment revenue for Apple
aapl_segments <- get_dimensional_data(
  metrics     = list("OperatingSegmentRevenue"),
  companies   = list("AAPL"),
  all_history = TRUE
)

# View results
aapl_segments[order(-fiscal_year, -fiscal_period),
              .(ticker, metric, value, container, fiscal_year, fiscal_period)]
```

---

## Search Disclosures and Footnotes

`get_disclosures()` searches SEC filing text sections: footnotes, MD&A, risk
factors, and other text blocks.

```{r get-disclosures, eval=has_creds}
# Risk factor disclosures for Apple, fiscal year 2023
aapl_disclosures <- get_disclosures(
  companies        = list("AAPL"),
  year             = 2023,
  disclosure_names = list("RiskFactors")
)
```

```{r disclosures-preview, eval=has_creds}
if (nrow(aapl_disclosures) > 0) {
  aapl_disclosures[, .(ticker, fiscal_year, disclosure_type_name, filing_date)]
}
```

### Full-text search

```{r disclosures-fts, eval=FALSE}
# Search all companies for "supply chain" mentions
supply_chain <- get_disclosures(
  companies             = list("AAPL", "MSFT"),
  full_text_search_term = "supply chain",
  all_history           = FALSE,
  year                  = 2023
)
```

---

## Get As-Reported Face Statements

`get_face_statement()` retrieves a single company's as-reported financial
statement exactly as filed with the SEC.

```{r get-face, eval=has_creds}
# Apple's annual income statement
aapl_income <- get_face_statement(
  company        = "AAPL",
  statement_type = "income",
  period_type    = "annual",
  all_history    = FALSE
)
```

```{r face-preview, eval=has_creds}
if (!is.null(aapl_income)) {
  cat("Statement:", aapl_income$name, "\n")
  cat("Columns:  ", length(aapl_income$columns), "\n")
  cat("Line items:", length(aapl_income$line_items), "\n")
}
```

Valid values for `statement_type`: `"income"`, `"balance"`, `"cash"`,
`"change-in-equity"`, `"comprehensive-income"`.

---

## Get Business Combinations (M&A)

`get_business_combinations()` returns purchase price allocation data for
acquisitions filed in SEC documents.

```{r get-bizcombo, eval=has_creds}
# Microsoft acquisitions
msft_acquisitions <- get_business_combinations(companies = list("MSFT"))
```

```{r bizcombo-preview, eval=has_creds}
if (nrow(msft_acquisitions) > 0) {
  msft_acquisitions[order(-acquisition_date),
                    .(parent_company_ticker, target, acquisition_date,
                      purchase_price)][1:5]
}
```

---

## Get Raw XBRL Data

`get_raw_xbrl()` retrieves individual XBRL data points as originally tagged in
SEC filings. Use `clauses` to filter by tag name, year, or period.

```{r get-raw-xbrl, eval=FALSE}
# All XBRL facts for Apple
aapl_xbrl <- get_raw_xbrl(companies = list("AAPL"))

# Filter by tag using a clause
revenue_xbrl <- get_raw_xbrl(
  companies = list("AAPL"),
  clauses   = list(
    list(value = "us-gaap:Revenues", parameter = "XBRLtag", operator = 1L)
  )
)
```

Clause operators: `1` = Equals, `10` = Contains, `21` = NotEquals.

---

## Get XBRL Tag Values (Multi-Company)

`get_xbrl_tag_values()` pulls standardized metric values for specific XBRL tags
or Calcbench metric names across multiple companies simultaneously.

```{r get-xbrl-tags, eval=FALSE}
# Revenue for S&P 500 via standardized metric name
revenue_all <- get_xbrl_tag_values(
  companies   = list("AAPL", "MSFT", "AMZN"),
  metrics     = list("Revenue"),
  all_history = FALSE
)

revenue_all[order(-fiscal_year, ticker),
            .(ticker, metric, value, fiscal_year, fiscal_period)]
```

---

## Error Handling

### 401 Unauthorized

A 401 error means one of two things:

1. **Wrong credentials** — check that `CALCBENCH_USERNAME` and `CALCBENCH_PASSWORD`
   are set correctly:
   ```{r check-env, eval=FALSE}
   Sys.getenv("CALCBENCH_USERNAME")
   Sys.getenv("CALCBENCH_PASSWORD")
   ```

2. **Insufficient subscription** — `mappedData` (standardized data) and
   `dimensionalData` (segments) require a premium account. Other endpoints
   (metrics, companies, filings) are available on all plans.
   Visit [calcbench.com/home/plans](https://www.calcbench.com/home/plans) to
   upgrade.

### Empty results

If a function returns an empty `data.table`, confirm:

- The company identifier matches Calcbench's name or ticker exactly
  (use `get_available_companies()` to look it up)
- The date/period range you specified contains filings
- Your account has access to the requested data type
